# Common Functions 

```{r include=FALSE}
library(tidyverse)
```


## For working with dataframes

I've scraped some data from r/psychologystudents so we can explore what we like to complain about. 

```{r include=FALSE}
psych_df <- read.csv("data/psychologystudents_20240820.csv")
```

```{r}
head(psych_df)
```

That's a bit chunky, why not we get rid of some variables? 

### select()

The select() function allows you retain specific columns (variables) by name from a dataframe. 

```{r}
psych_df %>%
  select(title) %>%
  head()
```

You can also pick out multiple columns by adding more arguments to the select() function.

```{r}
psych_df %>%
  select(score, id, subreddit) %>%
  head()
```

You can also retain columns by their numerical order. Here, the score, id, and subreddit are columns 3, 4, and 5 respectively.

```{r eval=FALSE}
psych_df %>%
  select(c(3:5)) %>%
  head()
```

### filter()

Recall that to call upon specific rows we can do this: 

```{r}
psych_df[4:7,] %>%
  select(title)
```

Which outputs rows 4 to 7.  

The filter function allows you to select specific rows that fit one or multiple criterion in your dataframe.

To get titles that have a Reddit score of 25 or more:

```{r}
psych_df %>%
  filter(score > 25) %>%
  select(title, score)

```

You can also input multiple logical conditions by adding more operators such as & or |. 
For a score of more than 25 and a "Discussion" flair. 

```{r}
psych_df %>%
  filter(score > 25 & flair == "Discussion") %>%
  select(title, score, flair) 
```

For a score of 100 or more OR a "Study Megathread" flair. 

```{r}
psych_df %>%
  filter(score >= 100 | flair == "Study Megathread") %>%
  select(title, score, flair)
```

### arrange()

Say I would like to look at 5 entries with the highest score. We can use arrange() and the function desc() to sort the text entries in a descending order:

```{r}
psych_df %>%
  select(title, score) %>%
  arrange(desc(score))%>%
  head(5)
```

5 lowest scoring entries? Arrange() sorts the dataframe in ascending order by default:

```{r}
psych_df %>%
  select(title, score) %>%
  arrange((score))%>%
  head(5)
```

Oh, seems like there might be more than 5 entries that scored a zero. Let's count how many using count():

```{r}
psych_df %>%
  filter(score==0) %>%
  count(score)
```

Seems like there are 8 entries with the score of 0.

### rbind()

This function is a super useful one. It binds rows of two or more dataframes together. Let's say I have these two dataframes:

```{r, include=FALSE}
name <- c("Bellatrix Lestrange","Basilisk","Voldemort","Barty Crouch Sr.","Professor Quirrell")
died <- c("Deathly Hallows","Chamber of Secrets","Deathly Hallows","Goblet of Fire","Philosopher's Stone")
bad_people <- data.frame(name,died)

name <- c("Cedric Diggory","Remus Lupin","Alastor Moody","Albus Dumbledore","Sirius Black")
died <- c("Goblet of Fire","Deathly Hallows","Deathly Hallows","Half Blood Prince","Order of the Phoenix")
good_people <- data.frame(name,died)
```

```{r}
good_people
```
```{r}
bad_people
```
Now it's time to use the rbind() function to bind the datasets together into one dataset. BUT, before we do that, it's good practice to add 


```{r}
hp_characters <- rbind(good_people,bad_people)
hp_characters
```
Great, so the rows of the second dataset (bad_people) were now added underneath the rows of the first dataset (good_people). But wait, what if someone doesn't know the characters that well so can't distinguish which characters are good or bad? It is good practice to add an extra column to each of the datasets beforehand so that you know where this data came from. Let's do that again:

```{r}
good_people <- good_people%>%
  mutate(identity="good")

bad_people <- bad_people%>%
  mutate(identity="bad")

hp_characters <- rbind(good_people,bad_people)
hp_characters
```

That's better.

### cbind()

This function is similar to rbind(), but instead of rows, it binds the columns (the c in cbind) of the datasets together. I have another dataset (actually just a column) that contains info about the house each of these characters are in:

```{r, include = FALSE}
house <- c("Hufflepuff","Gryffindoor","Unknown","Gryfindoor","Gryffindoor","Slytherin","NA","Slytherin","Unknown","Ravenclaw")

house_info <- data.frame(house)
```

```{r}
house_info
```

Let's add it to the hp_characters dataset.

```{r}
hp_characters_info <- cbind(hp_characters,house_info)
hp_characters_info
```

Note: Make sure that the number of rows match, otherwise you will get an error :/. 

Also, both the rows and the columns just get appended in the order that they are in. So make sure you double check that the data is as you want it to be.

### write.csv()

This function is good for when you want to save a dataset that you have created/cleaned in R to your computer as a csv file (which you can then put on OSF or something). This is the syntax:

```{r}
write.csv(hp_characters_info,"./data/filename.csv")
```

Note: The "./data/filename.csv" part of the argument specifies the path to the directory where you want to save it, ending with your chosen file name. If you don't specify the path, it will save it in your current directory.


## For working with text

Let's go back to the r/psychologystudents dataset and see what tools or working with text we can use.

### gsub()

This function let's you substitute a string or the characters in a vector or a dataframe with another string. It's very useful for data cleaning. For example, many big datasets of online comments usually contain a lot of links (starting with https:// ...). In most cases, we are not interested in those links, as they might interfere with our analyses (although sometimes they might also be interesting). There is a way to get rid of/replace things we are not interested in.

The syntax of gsub() is the following:

**gsub(pattern, replacement, x)**
  
  - pattern: The pattern or the string that you want to replace
  - replacement: A string you want to input instead of the pattern.
  - x: the dataset/column you want the replacement to happen


If we wanted to replace all hashtags with a blank space (basically just get rid of them) in the body column of our dataframe, we could do this:

```{r}
psych_df$body <- gsub("#","",psych_df$body)
```


It gets more complicated if you want to do more complex things. 

... 






