# Useful Packages

- don't forget that to install a package, you install.packages("tidyverse") and then run this to load the package:

```{r}
library(tidyverse)
```

Your good old friend:

## Tidyverse

is actually a package containing other packages that you will very likely use in most of your analyses. These core packages include:

  - ggplot2
  - dplyr
  - tidyr
  - readr
  - purrr
  - tibble
  - stringr
  - forcats
  
(but there's others too)
  
Let's have a look at some of the most useful ones.

### ggplot2

Great for plots! You have probably used it in DAPR already - you will use it in this course too. Usually, it starts with the function ggplot(), you then give it a dataset and specify what you want it to look like (aesthetic mapping) with aes(). Here is an example:

```{r,include=FALSE}
library(tidytext)

unnested <- psych_df %>%
  unnest_tokens(words,body)

data(stop_words)

stop_words <- stop_words %>%
  rename("words" = "word")

psych_words <- unnested %>%
  anti_join(stop_words)

psych_words <- psych_words %>%
  count(words, sort=TRUE)

psych_words <-  psych_words%>%
  filter(n > 30)%>%
  mutate(words = reorder(words,-n))

```

```{r}
ggplot(psych_words,aes(x=words,y=n))+
  geom_col(fill = "lightblue")+
  labs(title ="Most common words in the psych dataset")
```

Boring, but this is a plot that shows the most used words in our previous dataset (I did some data wrangling beforehand to create the psych_words dataframe - by tokenizing psych_df!).

Click here to find a nice Data Visualization with ggplot2 cheatsheet:

https://github.com/rstudio/cheatsheets/blob/main/data-visualization.pdf

