# Useful Packages

- don't forget that to install a package, you type install.packages("tidyverse") into the console and then run this to load the package:

```{r}
library(tidyverse)
```

First, your good old friend:

## Tidyverse

is actually a package containing other packages that you will very likely use in most of your analyses. These core packages include:

  - ggplot2
  - dplyr
  - tidyr
  - readr
  - purrr
  - tibble
  - stringr
  - forcats
  
(but there's others too)
  
Let's have a look at some of the most useful ones.

### ggplot2

Great for plots! You have probably used it in DAPR already - you will use it in this course too. Usually, it starts with the function ggplot(), you then give it a dataset and specify what you want it to look like (aesthetic mapping) with aes(). Here is an example:

```{r,include=FALSE}
library(tidytext)

unnested <- psych_df %>%
  unnest_tokens(words,body)

data(stop_words)

stop_words <- stop_words %>%
  rename("words" = "word")

psych_words <- unnested %>%
  anti_join(stop_words)

psych_words <- psych_words %>%
  count(words, sort=TRUE)

psych_words <-  psych_words%>%
  filter(n > 30)%>%
  mutate(words = reorder(words,-n))

```

```{r}
ggplot(psych_words,aes(x=words,y=n))+
  geom_col(fill = "lightblue")+
  labs(title ="Most common words in the psych dataset")
```

Boring, but this is a plot that shows the most used words in our previous dataset (I did some data wrangling beforehand to create the psych_words dataframe - by tokenizing psych_df!).

Click [here](https://github.com/rstudio/cheatsheets/blob/main/data-visualization.pdf) to find a nice Data Visualization with ggplot2 cheatsheet.

### dplyr

A very very useful package for data handling. The most common functions from the package are:

  - mutate() = adds a new variable (column) to the dataframe
  - select() = picks out variables from a dataframe based on their name
  - filter() = picks out cases based on a condition
  - summarise() = gives you a summary of multiple values
  - arrange() = arranges rows in a dataset (usually asceding or descending order)
  
Some other functions you might see Zach using are "join" functions, which help you combine dataframes and retain certain values from each, depending on what you need. They are a bit confusing, but hopefully this explanation helps:

I have two dataframes here:

```{r, include=FALSE}
course <- c('DAPR2', 'Psych in Action', 'Big Data', 'Psych 1A', 'DAPR3')
avg_sleep <- c('7', '10', '4', '9','6')
sleep <- data.frame(course, avg_sleep)
```

```{r}
sleep
```

```{r, include = FALSE}
course <- c('DAPR2', 'Psych in Action', 'Big Data', 'Psych 1A', 'Critical Analysis')
n_students <- c('250','163','39','300','161')
students <- data.frame(course,n_students)
```

```{r}
students
```

Now let's see what the join functions do:

**1. inner_join()**

- keeps observations from the first dataframe that have a matching key in the second dataframe (NOTE: join_by() specifies what variable should be our matching variable)

```{r}
inner_join(sleep, students, join_by(course))
```

This one put the two dataframes together but left out Critical Analysis and DAPR3, because they are only in one of the two dataframes.

**2. left_join()**

- keeps all observations in the first dataframe.

```{r}
left_join(sleep, students,join_by(course))
```

This time, Critical Analysis is left out because it is only in the second dataframe.

**3. right_join()**

- keeps all observations in the second dataframe

```{r}
right_join(sleep, students,join_by(course))
```

So, DAPR3 is left out because it is in the first dataframe.


**4. full_join()**

- keeps all observations from both dataframes

```{r}
full_join(sleep, students,join_by(course))
```

Now we've got all the observations, but some have got missing values.

**5. anti_join()**

- keeps all observations from the first dataframe without a match in the second one.

```{r}
anti_join(sleep, students ,join_by(course))
```
Just DAPR3 is left from the first dataframe.

**Different variable names**

Thus far, all these functions required the same variables names for the two datasets that you want to join. There are two methods that we can take to tackle joining variables of different names. 

```{r, include = FALSE}
courses <- c('DAPR2', 'Psych in Action', 'Big Data', 'Psych 1A', 'Critical Analysis')
satisfied <- c('strongly agree','moderate','somewhat disagree','agree','moderate')
satisfaction <- data.frame(courses, satisfied)
```

```{r}
satisfaction
```

1. Renaming the variable 

```{r}
satisfaction_renamed <- satisfaction %>%
  rename("course" = "courses")
satisfaction_renamed
```

```{r}
full_join(satisfaction_renamed, students,join_by(course))
```

2. Specifying which variables in each dataframe to compare

```{r}
full_join(satisfaction, students, by = c("courses" = "course"))
```

Here I specified the variable "courses" I want in the satisfaction dataframe and equate it to th variable "course" I want to compare to from the students dataframe.

<br>

Hopefully this was comprehensive enough for you to understand the bascis, but if you want more guidance check out this [cheatsheet](https://github.com/rstudio/cheatsheets/blob/main/data-transformation.pdf) for the dplyr package.



